\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-01-31 draft.tex
proof sec:pref

\subsection{Preferences}
\label{sec:pref}
I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare. This alters none of the results and so would only serve to complicate the exposition.}} The weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.

@@

In the first stage, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient. In this symmetric environment, it therefore maximizes the expected joint payoffs:
\begin{multline}
  {\color{blue}\expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right]} = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
  \label{eq:jv}
\end{multline}

@@

The home executive's welfare is specified as follows:
\[
  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
\]
Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort. This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).

@@

>>>>

\subsection{Preferences}
\label{sec:pref}
I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare. I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.

@@

In the first stage of the game, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient. In this symmetric environment, it therefore maximizes the expected joint payoffs:
\begin{multline}
  {\color{blue}\expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right]} = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
  \label{eq:jv}
\end{multline}

@@

The home executive's welfare is specified as follows:
\[
  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
\]
Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.

@@

This construction (exec vs. leg weight on X profits) permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).
